---
title: GC analysis - template
author: "Kenyon"
date: "February 1 2016"
output: html_document
---


#### GR analysis of CssS-7 set strains in DSM

Import files:

```{r echo=FALSE}
source('~/Bacillus_subtilis/Growth Rates/growth-curves-ryan/GrowthRateScripts_Dec2015/1_import_csvs.R')

folder_name<-"C:/Users/Kenyon/Documents/Bacillus_subtilis/Growth Rates/Eon data/DSM-LB-2rowseach+cssS/"
file_names<-list.files(folder_name,pattern="*[1-9].csv") #Fill me in
media<-"DSM"  #fill me in
today<-"2016-02-01"

obj_name<-"csvob"

file_names

```

Merge individual '.csv' files into one:
```{r echo=FALSE}
for (i in 1:length(file_names)){
  x<-read.csv(paste0(folder_name,file_names[i]),header = FALSE)
  assign(paste0(obj_name,i),x)
}

#Check to make sure length of this is correct:
csvob<-rbind(csvob1,csvob2[-1,],csvob3[-1,],csvob4[-1,],csvob5[-1,],csvob6[-1,],csvob7[-1,])

#csvob[,2]<-c("media",rep(media,times=nrow(csvob)-1))

#account for OD saturation:

nums<-csvob[-1,-1:-4]

csvob2<-(0.622*nums*nums)+(0.77*nums)

csvob3<-rbind(csvob[1,-1:-4],csvob2)
csvob<-cbind(csvob[,1:4],csvob3)

#media column already included

write.csv(csvob,paste0(folder_name,media,"_rawdatacollected_",today,".csv"))

#grab LB or DSM rows only (depending on name of "media" object)

csvob<-csvob[c(1,which(csvob[,2]==media)),]

```

* open grofit function, no overlay of plots so bad plots can be identified


```{r eval=FALSE}

source('~/Bacillus_subtilis/Growth Rates/growth-curves-ryan/no_overlay-gc_combined/grofit-combined-no_overlay.R')

par(mfrow=c(1,1))
gcdat<-gcfall_no(csvob)   #already run, do not need to do again

```


* all data to date processed fine, except that some curves have much flatter peaks while others are pointy

####observations: 


####data to remove: 

MN Oct-9 and Oct9 CN sample failed to grow, remove from set
Also CM Sept-23 and CN Sep-25

Also drop timpoints 80:89 - cause problems with peak & maxGR finding

```{r}
csvob[which(csvob[,70]<0.2),1:4]
csvob<-csvob[-which(csvob[,70]<0.2),c(1:80)]
```


using version of 'gcfall' that merges ln=TRUE and ln=FALSE datasets
```{r results='hide',message=FALSE,width=120,warning=FALSE}
source('~/Bacillus_subtilis/Growth Rates/growth-curves-ryan/GrowthRateScripts_Dec2015/3_gcfall-media-lnmerge.R')

gcdat2<-gcfall_mergeln(csvob)
```

**observations:**


#### save data and combine technical replicates
```{r echo=FALSE, width=120}
write.csv(gcdat2,paste0(folder_name,media,"-gcfit-",today,".csv"))

# Combine technical replicates
source('~/Bacillus_subtilis/Growth Rates/growth-curves-ryan/LM_controls-Sept2015/average_triples_ingcobj.R')

gcavg<-avg_triples(gcdat2,"DEF")

write.csv(gcavg,paste0(folder_name,media,"-gcfit-repsavgd",today,".csv"))

```


### plot of GR paramaters

```{r warning=FALSE,fig.width=11,fig.height=9}
source('~/Bacillus_subtilis/Growth Rates/growth-curves-ryan/LM_controls-Sept2015/gcf-tplot-by-strain.R')
errort.bars(gcavg)

```


**ln_mu.spline:**
nothing statistically sig dif, but all mutants except MN appear to be moving towards slower GR

**lambda.spline:**
all strains have significantly longer lag than AM373; also, CN is sig dif from NhaK

**A.spline:**
CssS sig higher than all other strains


    
```{r width=120}
source('~/Bacillus_subtilis/Growth Rates/growth-curves-ryan/LM_controls-Sept2015/gcftukeyrun_rM.R')

tuk_dat<-tukeyrun(gcavg,c(11,13:15))

tuk_dat[,3:5]<-signif(tuk_dat[,3:5],digits = 3)

print(tuk_dat[which(tuk_dat$pvals<0.05),])
print(tuk_dat)
```



