---
title: "cssS-DSM-epi-analysis_walkthrough"
author: "Kenyon"
date: "January 11, 2016"
output: html_document
---


### 1. Location of gcfile with growth rate measurements
```{r}

dat_name<-"DSM"
media<-"DSM"

gcfile<-read.csv("C:/Users/Kenyon/Documents/Bacillus_subtilis/Growth Rates/Eon data/DSM-LB-2rowseach+cssS/cssS-DSM-gcfit-data-avgdups.csv")
gcfile<-gcfile[,-1]
gcfile<-gcfile[-which(is.na(gcfile$strain)),]
```

### 2. Create stacked version of data:
Use "GRdata_collect_heirarch.R", which creates a stacked version of data  

Header of stacked version of data:
```{r echo = FALSE, warning=FALSE,message=FALSE,width=120}

options(width = 120)

name<-function(gcfile, med){
  library(dplyr)
  gcfile<-select(gcfile,plate,strain,well,log.y,plate.row,plate.col,lambda.spline,ln_mu.spline,A.spline,integral.spline)
  media<-rep(med,times=nrow(gcfile))
  
  gcfile<-cbind(media,gcfile)
  
  return(gcfile)
}

gc_named<-name(gcfile[,1:16],media)

hstack<-function(gcdat){
  
  library(dplyr)
  
  st.names<-levels(gcdat$strain)
  
  x3<-NULL
  
  for (i in 1:length(st.names))
  {
    x1<-gcdat[gcdat$strain == st.names[i],]
    
    x1<-select(x1,media,strain,plate,well,lambda.spline,ln_mu.spline,A.spline,integral.spline)
    
    x2<-stack(x1[,5:8])
    
    x2<-cbind(x2,x1$plate,x1$media,x1$well)
    colnames(x2)<-c(st.names[i],"assay","plate","media","well")
    
    x2<-x2[,c(2,4,3,5,1)]  # assay, media, plate, strain meas
    
    if(i==1){x3<-x2
    }else {x3<-cbind(x3,x2[match(paste0(x3[,1],"_",x3[,2],"_",x3[,3]),paste0(x2[,1],"_",x2[,2],"_",x2[,3])),5])}
    colnames(x3)[i+4]<-st.names[i]
  }
  
  #x3<-x3[-which(is.na(x3$plate)),]
  return(x3)
  
}

stack<-hstack(gc_named)
head(stack)

```

### 3. Plot of pre-normalized measurements, using errort.bars function:

```{r warning=FALSE,echo=FALSE,fig.height=8,fig.width=11}
options(width = 120)

errort.bars<-function(gcobj) {
  par(mfrow=c(2,2))
  
  gcobj$strain<-droplevels(gcobj$strain)
  gcobj$strain<-factor(gcobj$strain,levels=c("AM373","CssS","McpA","NhaK","CM","CN","MN"),ordered=TRUE)
  gcobj<-gcobj[order(gcobj$strain,gcobj$plate),]
  
  source('~/Bacillus_subtilis/Growth Rates/growth-curves-ryan/tplot.R')
  for (i in 8:11) {
    
tplot(gcobj[,i] ~ gcobj$strain, data=gcobj, las=1, cex=1, cex.axis=1, bty='L', show.n=F, dist=.25, jit=.05,
      type=c('db','db','db','db','db','db','db'),group.pch=T, pch= 6,
      col=c(1:length(levels(gcobj$plate)))[gcobj$plate], boxcol=c('lightsteelblue1'),
      boxborder=grey(.8),boxplot.pars=list(notch=F, boxwex=.5),main=paste(media," - ",colnames(gcobj)[i])[1])   
    abline(h=mean(gcobj[which(gcobj$strain=="AM373"),i]),col="red")
    }
}

errort.bars(gc_named)
print("units on y-axis of figures:  lambda.spline (hrs), ln_mu.spline (ln(OD)/hr), A.spline (OD600), integral.spline (OD*hrs)")

source('~/Bacillus_subtilis/Growth Rates/growth-curves-ryan/LM_controls-Sept2015/gcftukeyrun_rM.R')

tuk_data<-tukeyrun(gc_named,c(8:11))

tuk_data[,3:5]<-signif(tuk_data[,3:5],digits = 3)

print(tuk_data[which(tuk_data$pvals<0.1),])

```
feature      description
----------   ----------------------------------------------------
Lag          All KOs have longer lag than AM373 (~10%); CN has longest lag (similar to in Triton-X & LB)
Growth rate  All appear very similar; none significantly different (similar to LB & Triton-X)
A            All CssS-strains higher than all other strains (6%) - bigger dif than in LB or Triton-X 

 
### 4. Look for stat. sig. epistatic phenotypes
Following process in "epistasis_analysis-2015-12-07.R":

#### a. Normalize data:
  > in this document, all values are simply divided by the AM373 values because it is the only strain present in all sets; but if we are just looking at the TritonX set, then can divide by average within each day
  
Normalized data, header of table all replicates of all measurements:
  
```{r echo=FALSE,warning=FALSE,fig.width=11,fig.height=9}
options(width = 120)
avg<-apply(stack[,-1:-4],1,mean)
norm_dat<-cbind.data.frame(stack[,1:4],signif(stack[,-1:-4]/avg,4))
norm_dat<-norm_dat[,c(1:5,8,9,11,6,7,10)]
norm_dat$assay<-factor(norm_dat$assay,levels=c("lambda.spline","ln_mu.spline","A.spline","integral.spline"),ordered=TRUE)
head(norm_dat)
```

#### b. Mean and sd of each assay level:
```{r echo=FALSE,width=120,warning=FALSE}

dat_mean<-NULL
dat_sd<-NULL
assay<-NULL
media<-NULL
num<-NULL
    
levs<-levels(norm_dat$assay)
        
  for(i in 1:length(levs)){
    byassay<-norm_dat[which(norm_dat$assay==levs[i]),]  
          
    ex_means<-colMeans(byassay[,-1:-4],na.rm=TRUE)                 #calculate column means
    ex_sd<-apply(byassay[,-1:-4],2, function(x) sd(x,na.rm=TRUE))  #calculate column sds
          
    assay<-c(assay,as.character(byassay$assay[1]))
    media<-c(media,as.character(byassay$media[1]))
    num<-c(num,dim(byassay)[1])
          
    dat_mean<-rbind(dat_mean,ex_means)    #create mean table
    dat_sd<-rbind(dat_sd,ex_sd)         #create sd table
        }

d_mean<-cbind.data.frame(assay,media,num,round(dat_mean,digits=3) ) #table of all means of all strains in all assays
print("average value of all replicates of each assay & strain:")
print(d_mean)
d_sd<-cbind.data.frame(assay,media,num,dat_sd)    #table of all sds of all strains of all assays

prc_dif<-cbind(d_mean[,1:3],round((d_mean[,-1:-4]-d_mean$AM373)/d_mean$AM373,digits=3))
prc_difsd<-cbind(d_sd[,1:3],round(d_sd[,-1:-4]/d_mean$AM373,digits=3))
avg_sd<-round(rowMeans(d_sd[,-1:-4]/d_mean$AM373),digits=3)
prc_difsd<-cbind(prc_difsd,avg_sd)

print("table of percent differences of each mutant strain from AM373:")
prc_dif
print("table of st. dev. associated with each value in above table:")
prc_difsd

```

```{r echo=FALSE,message=FALSE,warning=FALSE,fig.height=8,fig.width=11}

#calculate normalized version of all measures, dividing each value by avg AM373 value for that assay

nrm<-arrange(norm_dat,assay)

AMrep<-rep(d_mean$AM373,each=7)
data<-apply(nrm[,5:11],2,function(x)(x-AMrep)/AMrep)

nrma<-cbind.data.frame(nrm[,1:4],data)
  
#generate unstacked version to use for plotting

library(tidyr)

stck<-gather(nrma,key=strain,value=measurement,AM373,CssS,McpA,NhaK,CM,CN,MN)
nrm2<-spread(stck,key=assay,value=measurement)
nrm3<-nrm2[,c(4,2,1,3,5:8)]


errort.bars2<-function(gcobj) {
  par(mfrow=c(2,2))
  
  gcobj$strain<-droplevels(gcobj$strain)
  gcobj$strain<-factor(gcobj$strain,levels=c("AM373","CssS","McpA","NhaK","CM","CN","MN"),ordered=TRUE)
  gcobj<-gcobj[order(gcobj$strain,gcobj$plate),]
  
  source('~/Bacillus_subtilis/Growth Rates/growth-curves-ryan/tplot.R')
  for (i in 5:8) {
    
tplot(as.numeric(gcobj[,i]) ~ gcobj$strain, data=gcobj, las=1, cex=1, cex.axis=1, bty='L', show.n=F, dist=.25, jit=.05,
      type=c('db','db','db','db','db','db','db'),group.pch=T, pch= 6,
      col=c(1:length(levels(gcobj$plate)))[gcobj$plate], boxcol=c('lightsteelblue1'),
      boxborder=grey(.8),boxplot.pars=list(notch=F, boxwex=.5),main=colnames(gcobj)[i])
     abline(h=0,col="red")}
}

errort.bars2(nrm3)

```

#### c.  Calculate estimated value of double mutants from the single mutants, and whether the measure phenotypes of the double mutants are sig dif from that estimate

Use function 'epi_func', not shown in report but provided in markdown file  

```{r echo=FALSE}
epi_func<-function(hstack,sing1,sing2,dbl){ #provide hstack of one media condition, data normalized, and names of desired strains
  
  mtab2<-hstack[,c("assay","media","AM373",sing1,sing2,dbl)]  #collects columns of desired strains
  mtab2<-mtab2[which(complete.cases(mtab2)),]
  
  #calculate percent change from AM373
  dat<-cbind.data.frame(mtab2[,1:2],(mtab2[,-1:-3]-mtab2$AM373)/mtab2$AM373)
  dbl<-((1+dat[,3])*(1+dat[,4]))-1
  dat<-cbind(dat,dbl)
  
  res<-NULL
  
  for(i in 1:length(levels(dat$assay))){
    dat1<-dat[which(dat$assay==levels(dat$assay)[i]),]
    pval<-t.test(dat1[,5],dat1[,6],paired = TRUE)$p.value
    w.pval<-wilcox.test(dat1[,5],dat1[,6],paired = TRUE)$p.value
    sig<-ifelse(pval<0.05|w.pval<0.05,"*","")
    
    res1<-c(levels(dat$assay)[i],as.character(dat$media[1]),colMeans(dat1[,3:6]),pval,w.pval,sig)
    res<-rbind(res,res1)
    
  }
  colnames(res)[7]<-"t.test"
  colnames(res)[8]<-"wil.test"
  
  colnames(res)[1:2]<-c("assay","media")
  row.names(res)[1:4]<-1:4
  
  res<-as.data.frame(res)
  
  res1<-apply(res[,3:8],2,function(x){
    signif(as.numeric(as.character(x)),digits=3)
  })
  
  res2<-cbind.data.frame(res[,1:2],res1,res[,9])
  colnames(res2)[9]<-"sig"
  
return(res2)
  }
```

##CM:
```{r warning=FALSE,echo=FALSE,fig.height=8,fig.width=11,width=120}

errort.bars2(nrm3[-which(nrm3$strain=="NhaK"|nrm3$strain=="CN"|nrm3$strain=="MN"),])
epi_func(norm_dat,"CssS","McpA","CM")

```

**CM: 2 Epistatic interactions found!** -

_Lag:_ Phenotype of double is between value of singles (+12% vs 10%+15%)
C & M have AM373 phenotype
CM's A.spline is significantly lower (2.3%) max OD than expected from single mutants (0.2%)   
**antagonistic epistasis** perhaps both KOs effect same mech, so functional loss redundant in double KO

_maxA:_ CM and McpA have ~AM373 (no change), while CssS has +2% peak maxOD **caution**: could be due to cssS-specific mutation, CM derived from McpA

##CN:
```{r warning=FALSE, echo=FALSE,fig.height=8,fig.width=11,width=120}
options(width = 120)
errort.bars2(nrm3[-which(nrm3$strain=="McpA"|nrm3$strain=="CM"|nrm3$strain=="MN"),])
epi_func(norm_dat,"CssS","NhaK","CN")
```

**CN: 1 Epistatic interaction, Lag** -

_Lag:_ Phenotype of CN is higher than both C & N, but not as high as additively predicted  **antagonistic epistasis** 


##MN:
```{r warning=FALSE, echo=FALSE,fig.height=8,fig.width=11,width=120}
options(width = 120)
errort.bars2(nrm3[-which(nrm3$strain=="CssS"|nrm3$strain=="CM"|nrm3$strain=="CN"),])
epi_func(norm_dat,"McpA","NhaK","MN")
```

**MN: 2 Epistatic interactions found!** -

_Lag:_ **antagonistic epistasis**: Phenotype of double is between value of singles (+11% vs 10%+12%)
M & N have AM373 phenotype

_maxA:_ **synergistic epistasis** Both M and N not sig lower (-1%) than AM373, while MN is -4% lower maxOD 

**why wasn't this noticed before?**: due to mutations acquired since big thaw?