---
title: "LB-TritonX GR analysis"
author: "Kenyon"
date: "December 22, 2015"
output: html_document
---


#### Grew 7 strains on LB+0.008% Triton X for 16hrs in plate reader, generating 7 replicates

Import files:

```{r eval=FALSE}
source('~/Bacillus_subtilis/Growth Rates/growth-curves-ryan/GrowthRateScripts_Dec2015/1_import_csvs.R')

folder_name<-"C:/Users/Kenyon/Documents/Bacillus_subtilis/Growth Rates/Eon data/LB+Triton-X/"
file_names<-list.files(folder_name,pattern="*tritonx...csv")
obj_name<-"tritX"

file_names

```

Merge individual '.csv' files into one:
```{r}
for (i in 1:length(file_names)){
  x<-read.csv(paste0(folder_name,file_names[i]),header = FALSE)
  assign(paste0(obj_name,i),x)
}

tritX<-rbind(tritX1,tritX2[-1,],tritX3[-1,],tritX4[-1,],tritX5[-1,],tritX6[-1,],tritX7[-1,])

tritX[,2]<-c("media",rep("LB+Trit",times=nrow(tritX)-1))

write.csv(tritX,"C:/Users/Kenyon/Documents/Bacillus_subtilis/Growth Rates/Eon data/LB+Triton-X/CMN-TritonXLB-2015-12-21.csv")

```

* open grofit function, no overlay of plots so bad plots can be identified

* in ln set, get 2 curves that are mis-identified at hr0; remove columns 5:15 (0.16-1.83hrs)

```{r eval=FALSE,width=120}
source('~/Bacillus_subtilis/Growth Rates/growth-curves-ryan/no_overlay-gc_combined/grofit-combined-no_overlay.R')

tridat<-gcfall_no(tritX)   #already run, do not need to do again

```


* all data to date processed fine, except that some curves have much flatter peaks while others are pointy

####observations:
* MN and AM373 generally had flat peaks, others were pointy

* the max point in flat peaks was usually identified at later timepoints in flat region

* appears as though CssS-containing strains lyse much faster than other strains

####data to remove:
* one MN growth curve appears much later than others; likely innoculation error
   + Dec-16, well D3
   
* several curves have expon. growth mis-identified in first 2hrs unless those timepoints ommitted (0.16-1.86hrs: data columns 5:15)

```{r}
tritX<-tritX[,-5:-15]
tritX<-tritX[-which(tritX[,1]=="Dec-16"& tritX[,4]=="D3"),]

```


using version of 'gcfall' that merges ln=TRUE and ln=FALSE datasets
```{r results='hide',message=FALSE,width=120,warning=FALSE}
source('~/Bacillus_subtilis/Growth Rates/growth-curves-ryan/GrowthRateScripts_Dec2015/3_gcfall-media-lnmerge.R')

tridat2<-gcfall_mergeln(tritX)
```

**observations:**

* non-CssS strains appear to decline into lysis less quicky (but not assayed as far into that region of growth)

* the max point in flat peaks was usually identified at later timepoints in flat region

* CssS strains appeared to enter exponential phase sooner, AM373 last (less/more lag)


#### save data and combine technical replicates:
```{r width=120}
write.csv(tridat2,"C:/Users/Kenyon/Documents/Bacillus_subtilis/Growth Rates/Eon data/LB+Triton-X/CMN-TritonXLB-gcfit-2015-12-22.csv")

# Combine technical replicates
source('~/Bacillus_subtilis/Growth Rates/growth-curves-ryan/LM_controls-Sept2015/average_triples_ingcobj.R')

trit<-avg_triples(tridat2,"DEF")

write.csv(trit,"C:/Users/Kenyon/Documents/Bacillus_subtilis/Growth Rates/Eon data/LB+Triton-X/CMN-TritonXLB-gcfit-repsavgd-2015-12-21.csv")

```


### plot of GR paramaters

```{r warning=FALSE,fig.width=11,fig.height=9}
source('~/Bacillus_subtilis/Growth Rates/growth-curves-ryan/LM_controls-Sept2015/gcf-tplot-by-strain.R')
errort.bars(trit)

```



**ln_mu.spline:**

    * CM is only stat dif from AM373 (8% higher); is also seg. from N+MN
    * MN is lowest, seg from all but McpA/NhaK - may still be epistatic though
    * MN looks like it has reduced growth compared to what is attributable to M/N
    * CM looks like it might have higher GR compared to C/M

**lambda.spline:**

    * all CssS mutants look like they have similarly reduced (6-10%) lag, while A,M,N,MN are
    * largely or only minorly different;  TukeyHSD completely backs this clustering
    
```{r width=120}
source('~/Bacillus_subtilis/Growth Rates/growth-curves-ryan/LM_controls-Sept2015/gcftukeyrun_rM.R')

tri_tuk<-tukeyrun(trit,c(11,13:15))

tri_tuk[,3:5]<-signif(tri_tuk[,3:5],digits = 3)

print(tri_tuk[which(tri_tuk$pvals<0.05),])
print(tri_tuk)
```



