---
title: "TritonX-epistasis analysis"
author: "Kenyon"
date: "January 21, 2016"
output: html_document
---



## In this document, I am analyzing the Triton-X 0.008% growth rate data for evidence of epistatic interactions

### In this version of this document, I am adding data from CM & CN clones


### 1. Location of gcfile with growth rate measurements
```{r}
tritX_gcfile1<-read.csv("C:/Users/Kenyon/Documents/Bacillus_subtilis/Growth Rates/Eon data/LB+Triton-X/CMN-TritonXLB-gcfit-repsavgd-2015-12-21.csv")

CM<-read.csv("C:/Users/Kenyon/Documents/Bacillus_subtilis/Growth Rates/Eon data/CMCNclones-TritonX/LB+TritCM-gcfit-repsavgd2016-01-21.csv")

CM<-CM[-which(CM$strain=="McpA_K"|CM$strain=="CM4"),]

CN<-read.csv( "C:/Users/Kenyon/Documents/Bacillus_subtilis/Growth Rates/Eon data/CMCNclones-TritonX/LB+TritCN-gcfit-repsavgd2016-01-21.csv")

CN<-CN[-which(CN$strain=="CN5"),]

#MN double mutants data was in LB, not LB-TritonX, so cannot combine with this data

tritX_gcfile<-rbind(tritX_gcfile1,CM,CN)
```

####1.b - correct A.spline data

```{r echo=FALSE}

ODmax<-tritX_gcfile$A.spline
ODmax_cor<-(0.399*ODmax*ODmax)+(0.873*ODmax)
tritX_gcfile$A.spline<-ODmax_cor
```
### 2. Create stacked version of data:

Header of stacked version of data:
```{r echo = FALSE, warning=FALSE,message=FALSE,width=120}

options(width = 120)

name<-function(gcfile, med){
  library(dplyr)
  gcfile<-select(gcfile,plate,strain,well,log.y,plate.row,plate.col,lambda.spline,ln_mu.spline,A.spline,integral.spline)
  media<-rep(med,times=dim(gcfile)[1])
  
  gcfile<-cbind(media,gcfile)
  
  return(gcfile)
}

tritX<-name(tritX_gcfile[,1:16],"TritX")

hstack<-function(gcdat){
  
  library(dplyr)
  
  gcdat$strain<-droplevels(gcdat$strain)
  
  st.names<-levels(gcdat$strain)
  
  x3<-NULL

  
  for (i in 1:length(st.names))
  {
    x1<-gcdat[gcdat$strain == st.names[i],]
    
    x1<-select(x1,media,strain,plate,well,lambda.spline,ln_mu.spline,A.spline,integral.spline)
    
    x2<-stack(x1[,5:8])
    
    x2<-cbind(x2,x1$plate,x1$media,x1$well)
    colnames(x2)<-c(st.names[i],"assay","plate","media","well")
    
    x2<-x2[,c(2,4,3,5,1)]  # assay, media, plate, strain meas
    
    
    #if AM373 set, then assign to x3; if not, find and match rows according to datalabels in AM373 set
    
    if(i==1){x3<-x2
    x3col<-substr(x3[,4],1,2)
    }else {x3<-cbind(x3,x2[match(paste0(x3[,1],"_",x3[,2],"_",x3[,3],"_",substr(x3[,4],1,2)),paste0(x2[,1],"_",x2[,2],"_",x2[,3],"_",substr(x2[,4],1,2))),5])}
    colnames(x3)[i+4]<-st.names[i]
  }
  
  #x3<-x3[-which(is.na(x3$plate)),]
  return(x3)
  
}

tritXb<-hstack(tritX)
head(tritXb)

```


### 3. Plot of pre-normalized measurements, using errort.bars function:

```{r warning=FALSE,echo=FALSE,fig.height=6,fig.width=8}

source('~/Bacillus_subtilis/Growth Rates/growth-curves-ryan/LM_controls-Sept2015/gcftukeyrun_rM.R')
tri_tuk<-tukeyrun(tritX,c(8:11))
tri_tuk[,3:5]<-signif(tri_tuk[,3:5],digits = 3)

notes<-NULL

notes[8]<-"lambda.spline: all strains but MN have shorter lag, especially CssS strains"
notes[9]<-"ln_mu.spline: CssS strains all have stat sig higher GR, 4-6%"
notes[10]<-"A.spline: only CM and CN are stat dif from AM373, and all other strains"


errort.bars<-function(gcobj) {
  #par(mfrow=c(2,2))
  
  gcobj$strain<-droplevels(gcobj$strain)
  gcobj$strain<-factor(gcobj$strain,levels=c("AM373","CssS","McpA","NhaK","CM","CN","MN"),ordered=TRUE)
  gcobj<-gcobj[order(gcobj$strain,gcobj$plate),]
  
  source('~/Bacillus_subtilis/Growth Rates/growth-curves-ryan/tplot.R')
  for (i in 8:10) {
    
tplot(gcobj[,i] ~ gcobj$strain, data=gcobj, las=1, cex=1, cex.axis=1, bty='L', show.n=F, dist=.25, jit=.05,
      type=c('db','db','db','db','db','db','db'),group.pch=T, pch= 6,
      col=c(1:length(levels(gcobj$plate)))[gcobj$plate], boxcol=c('lightsteelblue1'),
      boxborder=grey(.8),boxplot.pars=list(notch=F, boxwex=.5),main=colnames(gcobj)[i]) 
    
      abline(h=mean(gcobj[which(gcobj$strain=="AM373"),i]),col="red") 
      
      print(notes[i])
      
      print(tri_tuk[which(tri_tuk$pvals<0.05 & tri_tuk$meas==colnames(gcobj)[i]),])
      }
}

errort.bars(tritX)
```

* The 2 last assays performed used new LB+TritonX media; has distinct lag & GR phenotype, but seems to trend similarly across strains

* CssS mutants all appear to have reduced lag and increased ln.mu compared to AM373 / other strains  
* CM/CN appear to have reduced A.spline compared to all other strains  
* all mutants appear to have increased integrated AUC compared to AM373; integral partitions into 2 groups, likely Jan vs Dec groups given lack of partition in MN; Jan integral appears higher - did I grow these for 2 hours longer? 


### 4.Quantify epistatic interactions


#### a. Normalize data:
  > in this set, all assays only have AM373 and CssS in common
  
Normalized data, header of table all replicates of all measurements:

* because of CM/CN data, no longer have all strains in all assays
* now will normalize by average of strains that are in all measurements - A & C

```{r echo=FALSE,warning=FALSE}
avg<-((tritXb$AM373+tritXb$CssS)/2)
trit_nm<-cbind.data.frame(tritXb[,1:4],signif(tritXb[,-1:-4]/avg,4))
trit_nm<-trit_nm[,c(1:5,8,9,11,6,7,10)]
trit_nm$assay<-factor(trit_nm$assay,levels=c("lambda.spline","ln_mu.spline","A.spline","integral.spline"),ordered=TRUE)
```

#### b. Mean and sd of each assay level:
```{r echo=FALSE,width=120,warning=FALSE}

dat_mean<-NULL
dat_sd<-NULL
assay<-NULL
media<-NULL
num<-NULL
    
levs<-levels(trit_nm$assay)
        
  for(i in 1:length(levs)){
    byassay<-trit_nm[which(trit_nm$assay==levs[i]),]  
          
    ex_means<-colMeans(byassay[,-1:-4],na.rm=TRUE)                 #calculate strain means
    ex_sd<-apply(byassay[,-1:-4],2, function(x) sd(x,na.rm=TRUE))  #calculate strain sds
          
    assay<-c(assay,as.character(byassay$assay[1]))
    media<-c(media,as.character(byassay$media[1]))
    num<-c(num,nrow(byassay))
          
    dat_mean<-rbind(dat_mean,ex_means)    #create mean table
    dat_sd<-rbind(dat_sd,ex_sd)         #create sd table
        }

d_mean<-cbind.data.frame(assay,media,num,round(dat_mean,digits=3) ) #table of all means of all strains in all assays


```

**average value of all replicates of each assay & strain:**

```{r echo=FALSE,width=120,warning=FALSE}
print(d_mean)
d_sd<-cbind.data.frame(assay,media,num,dat_sd)    #table of all sds of all strains of all assays

#calculate percent difference of each strain's average value relative to AM373's average
prc_dif<-cbind(d_mean[,1:3],round((d_mean[,-1:-4]-d_mean$AM373)/d_mean$AM373,digits=3))

#calculate sd of each, relative to %AM373 value
prc_difsd<-cbind(d_sd[,1:3],round(d_sd[,-1:-4]/d_mean$AM373,digits=3))
avg_sd<-round(rowMeans(d_sd[,-1:-4]/d_mean$AM373),digits=3)


#add avg_sd to prc_dif to make comparing effect sizes to avg_sd easier
prc_dif<-cbind(prc_dif,avg_sd)

```

**table of percent differences of each mutant strain from AM373:**

```{r echo=FALSE,width=120,warning=FALSE}
prc_dif
```

**table of st. dev. associated with each value in above table:**

```{r echo=FALSE,width=120,warning=FALSE}
prc_difsd
```

####Plots of normalized data:
* Noramlized by (x-avgAM373)/avgAM373

```{r echo=FALSE,message=FALSE,warning=FALSE,fig.height=8,fig.width=11}

#calculate normalized version of all measures, dividing each value by avg AM373 value for that assay

nrm<-arrange(trit_nm,assay)

AMrep<-rep(d_mean$AM373,each=table(nrm$assay)[1])
data<-apply(nrm[,5:11],2,function(x)(x-AMrep)/AMrep)

nrma<-cbind.data.frame(nrm[,1:4],data)
  
#generate unstacked version to use for plotting

library(tidyr)

stck<-gather(nrma,key=strain,value=measurement,AM373,CssS,McpA,NhaK,CM,CN,MN)
nrm2<-spread(stck,key=assay,value=measurement)
nrm3<-nrm2[,c(4,2,1,3,5:8)]


errort.bars2<-function(gcobj) {
  par(mfrow=c(2,2))
  
  gcobj$strain<-droplevels(gcobj$strain)
  gcobj$strain<-factor(gcobj$strain,levels=c("AM373","CssS","McpA","NhaK","CM","CN","MN"),ordered=TRUE)
  gcobj<-gcobj[order(gcobj$strain,gcobj$plate),]
  
  source('~/Bacillus_subtilis/Growth Rates/growth-curves-ryan/tplot.R')
  for (i in 5:8) {
    
tplot(as.numeric(gcobj[,i]) ~ gcobj$strain, data=gcobj, las=1, cex=1, cex.axis=1, bty='L', show.n=F, dist=.25, jit=.05,
      type=c('db','db','db','db','db','db','db'),group.pch=T, pch= 6,
      col=c(1:length(levels(gcobj$plate)))[gcobj$plate], boxcol=c('lightsteelblue1'),
      boxborder=grey(.8),boxplot.pars=list(notch=F, boxwex=.5),main=colnames(gcobj)[i])
    abline(h=0,col="red")}
}

errort.bars2(nrm3)

```


#### c.  Calculate estimated value of double mutants from the single mutants, and whether the measure phenotypes of the double mutants are sig dif from that estimate

1. normalize each assay by the AM373 measurement (x-AM373)/AM373
2. estimate double-mutant value by adding effect size of single mutants
3. perform t.test and wilcoxon rank sum test to compare measured and estimated values, estimate if sig dif
 + using paired=TRUE
 
 + also, plots include all data, while epistasis calculations only include complete cases - so AM373 & CssS plots show datapoints not included in calculation

```{r echo=FALSE}
epi_func<-function(hstack,sing1,sing2,dbl,type){ #provide hstack of one media condition, data normalized, and names of desired strains
  
  mtab2<-hstack[,c("assay","media","AM373",sing1,sing2,dbl)]  #collects columns of desired strains
  mtab2<-mtab2[which(complete.cases(mtab2)),]
  reps<-mean(table(mtab2$assay)) #get number of assays used in calculations
  
  #calculate percent change from AM373
  dat<-cbind.data.frame(mtab2[,1:2],(mtab2[,-1:-3]-mtab2$AM373)/mtab2$AM373)  #get %change of each meas. from AM373 
  
  if(type=="mult"){
    dbl_est<-((1+dat[,3])*(1+dat[,4]))-1} else if(type=="add"){
      dbl_est<-dat[,3]+dat[,4]}else{
        print("epi_func var type not specified - either add or mult")}
  
  dat<-cbind(dat,dbl_est)
  
  res<-NULL
  
  for(i in 1:length(levels(dat$assay))){
    dat1<-dat[which(dat$assay==levels(dat$assay)[i]),]
    pval<-t.test(dat1[,5],dat1[,6],paired = TRUE)$p.value #dat[,5] = actual dbl meas.; dat[,6]= estim dbl meas.
    w.pval<-wilcox.test(dat1[,5],dat1[,6],paired = TRUE)$p.value
    sig<-ifelse(pval<0.05|w.pval<0.05,"*","")
    
    res1<-c(levels(dat$assay)[i],as.character(dat$media[1]),colMeans(dat1[,3:6]),pval,w.pval,sig)
    res<-rbind(res,res1)
    
  }
  colnames(res)[7]<-"t.test"
  colnames(res)[8]<-"wil.test"
  
  colnames(res)[1:2]<-c("assay","media")
  row.names(res)[1:4]<-1:4
  
  res<-as.data.frame(res)
  
  res1<-apply(res[,3:8],2,function(x){
    signif(as.numeric(as.character(x)),digits=3)
  })
  
  res2<-cbind.data.frame(res[,1:2],reps,res1,res[,9])
  colnames(res2)[10]<-"sig"
  
return(res2)
  }


```



##CM:
```{r warning=FALSE,echo=FALSE,fig.height=8,fig.width=11,width=120}

errort.bars2(nrm3[-which(nrm3$strain=="NhaK"|nrm3$strain=="CN"|nrm3$strain=="MN"),])
epi_func(trit_nm,"CssS","McpA","CM",type="mult")
epi_func(trit_nm,"CssS","McpA","CM",type="add")
```

Epistatic interaction found! - on A. spline
C & M have AM373 phenotype
CM's A.spline is significantly lower (2.3%) max OD than expected from single mutants (0.2%)

also has less gain of integrated area (+5%) than expected from single mutants (7.7%+3.3% = 11% gain expected)
  - instead, expresses intermediate value between the two single mutants

##CN:
```{r warning=FALSE, echo=FALSE,fig.height=8,fig.width=11,width=120}
options(width = 120)
errort.bars2(nrm3[-which(nrm3$strain=="McpA"|nrm3$strain=="CM"|nrm3$strain=="MN"),])
epi_func(trit_nm,"CssS","NhaK","CN",type="mult")
epi_func(trit_nm,"CssS","NhaK","CN",type="add")
```
same with CN -
Epistatic interaction found! - on A. spline
C & N have AM373 phenotype (-0.0)
CM's A.spline is significantly lower (1.7%) max OD than expected from single mutants (0.3%)


##MN:
```{r warning=FALSE, echo=FALSE,fig.height=8,fig.width=11,width=120}
options(width = 120)
errort.bars2(nrm3[-which(nrm3$strain=="CssS"|nrm3$strain=="CM"|nrm3$strain=="CN"),])
epi_func(trit_nm,"McpA","NhaK","MN",type="mult")
epi_func(trit_nm,"McpA","NhaK","MN",type="add")
```

No epistatic effects found in MN mutant
