---
title: "Epistasis analysis of CssS-recovered set in LB-pH5 media"
author: "Kenyon"
date: "June 25, 2016"
output: html_document
---


### Describe dataset  

Regenerated CssS strain set grown on LB acidified to pH5 with HCl

Strain set consists of single-gene deletion strains:  
     
* AM373: Reference strain; no deletions  
* [cssS:](http://subtiwiki.uni-goettingen.de/bank/index.php?gene=cssS&action=Go) two-component sensor kinase gene that response to secretion stress deleted    
* [mcpA:](http://subtiwiki.uni-goettingen.de/bank/index.php?gene=mcpA&action=Go) membrane-bound chemotaxis receptor gene deleted  
* [nhaK:](http://subtiwiki.uni-goettingen.de/bank/index.php?gene=nhaK&action=Go) cation/H+ antiporter gene deleted  

Double-gene deletion strains:  
    
* CM : cssS/mcpA double KO strain  
* CN : cssS/nhaK KO  
* MN : mcpA/nhaK KO  

### 1. Import growth rate measurement files


```{r warning=FALSE,message=FALSE}

#Import growth curve data file:
gcfile<-read.csv("C:/Users/Kenyon/Documents/Bacillus_subtilis/Growth Rates/Eon data/LB+HCL-PH5/LB_pH5-gcfit-repsavgd2016-06-14.csv",row.names=1)
gcfile<-gcfile[,-1]

#text variables:
media<-"LB_pH5"
strain_order<-c("AM373","CssS","McpA","NhaK","CM","CN","MN")

library(knitr)
library(dplyr)
library(tidyr)

```



### 2. Reshape to create stacked version of data
 
```{r echo = FALSE, warning=FALSE,message=FALSE,width=120}
#grab only needed columns
meascol<-select(gcfile,plate,strain,ln_mu.spline,mu.spline,lambda.spline,A.spline,integral.spline)

stack<-gather(meascol,key=assay,value = measurement,ln_mu.spline,mu.spline,lambda.spline,A.spline,integral.spline )

strcol<-spread(stack,key=strain,value=measurement)

```

### 3. Plots of raw measurements by strain:    
    
* red line indicates AM373-average in assay   
* ln_mu.spline :  max growth rate of ln-transformed data (doublings/hr)  
* mu.spline: max growth rate, raw data (delta-OD/hr)  
* lambda.spline:  lag time (hrs)  
* A.spline: MaxOD of growth curve  (OD)  
* integral.spline: integrated area under each growth curve

```{r warning=FALSE,echo=FALSE,fig.height=8,fig.width=11}
options(width = 120)

errort.bars<-function(gcobj) {
  layout(matrix(c(1,2,3,4,5,6), ncol=2, byrow=TRUE), heights=c(2,2,2))
  
  gcobj$strain<-droplevels(gcobj$strain)
  gcobj$strain<-factor(gcobj$strain,levels=strain_order,ordered=TRUE)
  gcobj<-gcobj[order(gcobj$strain,gcobj$plate),]
  
  source('~/Bacillus_subtilis/Growth Rates/growth-curves-ryan/tplot.R')
  for (i in 3:7) {
    
tplot(gcobj[,i] ~ gcobj$strain, data=gcobj, las=1, cex=1, cex.axis=1, bty='L', show.n=F, dist=.25, jit=.05,
      type=rep('db',times=length(levels(gcobj$strain))),group.pch=T, pch= rep(6,times=length(levels(gcobj$strain))),
      col=c(1:length(levels(gcobj$plate)))[gcobj$plate], boxcol=c('lightsteelblue1'),
      boxborder=grey(.8),boxplot.pars=list(notch=F, boxwex=.5),main=paste(media," - ",colnames(gcobj)[i])[1])   
    abline(h=mean(gcobj[which(gcobj$strain=="AM373"),i]),col="red")
  }
  
  par(mai=c(0.5,0.5,0.5,0.5))
  frame()
  legend(x="bottom",
         legend=levels(gcobj$plate),pch=6,col=c(1:length(levels(gcobj$plate)))[gcobj$plate],
         title="date grown")
  
  }

errort.bars(meascol)
print("units on y-axis of figures:  lambda.spline (hrs), ln_mu.spline (ln(OD)/hr), A.spline (OD600), integral.spline (OD*hrs)")

source('~/Bacillus_subtilis/Growth Rates/growth-curves-ryan/LM_controls-Sept2015/gcftukeyrun_rM.R')

tuk_data<-tukeyrun(meascol,c(3:7))

tuk_data[,3:5]<-signif(tuk_data[,3:5],digits = 3)

#print(tuk_data[which(tuk_data$pvals<0.1),])

```

 
### 4. Normalization


#### a. Normalize data by experiment (plate) average

  
```{r echo=FALSE,warning=FALSE,fig.width=11,fig.height=9}

avg<-apply(strcol[,-1:-2],1,mean) #calculate row average
norm_dat<-cbind.data.frame(strcol[,1:2],signif(strcol[,-1:-2]/avg,4))  #divide each measurement by its avg daily value across strains
norm_dat<-norm_dat[,c(1,2,3,6,7,9,4,5,8)] #rearrange columns to get strains in order "A,C,M,N,CM,CN,MN"
norm_dat$assay<-factor(norm_dat$assay,levels=c("lambda.spline","A.spline","ln_mu.spline","mu.spline", "integral.spline"),ordered=TRUE)

```


**average normalized values of each assay, by strain:**

```{r echo=FALSE,width=120,warning=FALSE}

dat_mean<-NULL
dat_sd<-NULL
assay<-NULL
num<-NULL
power<-NULL
    
levs<-levels(norm_dat$assay)
        
  for(i in 1:length(levs)){
    byassay<-norm_dat[which(norm_dat$assay==levs[i]),]  
          
    ex_means<-colMeans(byassay[,-1:-2],na.rm=TRUE)                 #calculate column means
    ex_sd<-apply(byassay[,-1:-2],2, function(x) sd(x,na.rm=TRUE))  #calculate column sds
          
    assay<-c(assay,as.character(byassay$assay[1]))
    num<-c(num,dim(byassay)[1])
    
    #power calculated using the average of the standard deviations across the strains
    power1<-power.t.test(n=num[i],sd=mean(ex_sd),sig.level=0.05,power=0.8,type="two.sample")$delta
    power<-signif(c(power,power1),digits=3)

    dat_mean<-rbind(dat_mean,ex_means)    #create mean table
    dat_sd<-rbind(dat_sd,ex_sd)         #create sd table
        }

d_mean<-cbind.data.frame(assay,num,round(dat_mean,digits=3)) #table of all means of all strains in all assays
d_sd<-cbind.data.frame(assay,num,round(dat_sd,digits=3))    #table of all sds of all strains of all assays

kable(cbind(d_mean,power))

```

**%-differences of each strain from AM373:**

```{r echo=FALSE,width=120,warning=FALSE}
prc_dif<-cbind(d_mean[,1:2],round((d_mean[,-1:-3]-d_mean$AM373)/d_mean$AM373,digits=3))

prc_difsd<-cbind(d_sd[,1:2],round(d_sd[,-1:-3]/d_mean$AM373,digits=3))

avg_sd<-round(rowMeans(d_sd[,-1:-2]/d_mean$AM373),digits=3)
prc_difsd<-cbind(prc_difsd,avg_sd)

power<-signif(apply(prc_difsd,1,function(x){power.t.test(n=as.numeric(x[2]),sd=mean(as.numeric(x[3:8])),sig.level=0.05,power=0.8,type="paired")$delta}),digits=2)

prc_difx<-cbind(prc_dif,avg_sd,power)
kable(prc_difx)
```

**%-differences, if above experiment-power:**

```{r echo=FALSE,width=120,warning=FALSE}
#generate a table with only entries if greater than power level
prc_table<-apply(prc_difx,1,function(x){
  
  y<-x[3:8]
  
  y[abs(as.numeric(y))<as.numeric(x[10])]<-NA
  x[3:8]<-y
  x[is.na(x)]<-as.character(" ")
  return(x)
})

prc_table<-t(prc_table)

kable(prc_table[,-9])


```

* power calculation is based on alpha=0.05, beta=0.2, paired measurements
  

#### b. Normalized all measurements by dividing each value by avg AM373 value for that measurement

After normalization, a new parameter is also generated - euclidian distance across the growth rate features
  `eucld_dist = sqrt((lambda.spline^2)+(A.spline^2)+(mu.spline^2)+(ln_mu.spline^2)+(integral.spline^2))`

Estimated phenotype of the double mutants is also calculated at this stage, named "XX_est", and plotted in orange next to the measured phenotype of each double mutant (dark blue)  
  
The est-double mutant phenotype is the sum of the effects of the single mutants in each replicate experiment -  
ie, `(sing1 - AM373) + (sing2 - AM373)`



```{r echo=FALSE,message=FALSE,warning=FALSE,fig.height=8,fig.width=11}

nrm<-arrange(norm_dat,assay)  #need assays to be sorted in same order as in d_mean

Ameans<-tapply(norm_dat$AM373,norm_dat$assay,mean)
AMrep<-rep(Ameans,each=length(levels(gcfile$plate)))
    
#calculate distance between each rep and the avg_AM373 value (not the set's AM373 value)
data<-apply(nrm[,3:9],2,function(x)(x-AMrep)/AMrep)

nrma<-cbind.data.frame(nrm[,1:2],data)  # %change from avgAM373, columns are strains

#change to assay-column format to calculate euclidian distance

nrm_a1<-nrma %>% gather(key=strain,value=measurement,AM373,CssS,McpA,NhaK,CM,CN,MN) %>%
  spread(key=assay,value=measurement) %>% 
  mutate(eucld_dist = sqrt((lambda.spline^2)+(A.spline^2)+(mu.spline^2)+(ln_mu.spline^2)+(integral.spline^2)))

#calculate estimated double-mutants here (additive model)
nrm_s<- nrm_a1 %>% gather(key=assay,value=measurement,lambda.spline,A.spline,ln_mu.spline,mu.spline,integral.spline,eucld_dist) %>% 
  spread(key=strain,value=measurement) %>%
  mutate(CM_est = (CssS-AM373)+(McpA-AM373),
       CN_est = (CssS-AM373)+(NhaK-AM373),
       MN_est = (McpA-AM373)+(NhaK-AM373))
  

#generate unstacked version to use for plotting
nrm_a<-nrm_s %>% gather(key=strain,value=measurement,AM373,CssS,McpA,NhaK,CM,CN,MN,CM_est,CN_est,MN_est) %>% spread(key=assay,value=measurement)


```

```{r echo=FALSE,message=FALSE,warning=FALSE,fig.height=8,fig.width=11}
errort.bars2<-function(gcobj) {
  par(mfrow=c(3,2))
  
  gcobj$strain<-droplevels(gcobj$strain)
  gcobj$strain<-factor(gcobj$strain,levels=c("AM373","CssS","McpA","NhaK","CM","CM_est","CN","CN_est","MN","MN_est"),ordered=TRUE)
  gcobj<-gcobj[order(gcobj$strain,gcobj$plate),]
  
  source('~/Bacillus_subtilis/Growth Rates/growth-curves-ryan/tplot.R')
  for (i in 3:8) {
    
tplot(as.numeric(gcobj[,i]) ~ gcobj$strain, data=gcobj, las=1, cex=1, cex.axis=1, 
      bty='L', show.n=F, dist=.25, jit=.05,
      type=rep('db',times=length(levels(gcobj$strain))),group.pch=T, 
      pch= rep(6,times=length(levels(gcobj$strain))),
      col=c(1:length(levels(gcobj$plate)))[gcobj$plate], 
      boxcol=c('azure3',rep('lightblue',times=3),rep(c('blue','coral'),times=3)),
      boxborder=grey(.8),boxplot.pars=list(notch=F, boxwex=.5),main=colnames(gcobj)[i])
     abline(h=0,col="red")
     
     }
}

errort.bars2(nrm_a)
#write.csv(nrm3,"~/Bacillus_subtilis/Growth Rates/Eon data/LB+HCL-PH5/CsetREGEN_pH5_normalized.csv")

```

### 5. Identify Epistatic Interactions

Calculate estimated value of double mutants from the single mutants, and whether the measure phenotypes of the double mutants are sig dif from that estimate 

In the t-box plots, looking for features in which the measured double-mutant phenotype (dark blue) is well-separated from the estimated phenotype (orange)

__Two Epistasis Calculation Methods:__

__Linear Modeling Results:__ Identify epistasis using linear modeling to identify significant interaction effects
    * in printed output, the double-mutant column (ie, "CM") is not the % dif from AM373-avg, but is the difference between double-mutant observed and double-estimated phenotype (ie, CM - CMest in the second table)  
    
 __Estimated Phenotype Calculation Results:__ Calcuate epistasis using t.test between observed and estimated double-mutant phenotype  



```{r echo=FALSE}
#input: norm_dat, which is normalized by plate average but not by AM373 yet, and has strains as columns

epi_func<-function(df,sing1,sing2,dbl,dbl_est,media){ #provide hstack of one media condition, data normalized, and names of desired strains
  
  mtab2<-df[,c("assay","AM373",sing1,sing2,dbl,dbl_est)]  #collects columns of desired strains
  dat<-mtab2[which(complete.cases(mtab2)),]
  
  res<-NULL
  
  for(i in 1:length(levels(dat$assay))){
    dat1<-dat[which(dat$assay==levels(dat$assay)[i]),]
    pval<-t.test(dat1[,5],dat1[,6],paired = TRUE)$p.value
    w.pval<-wilcox.test(dat1[,5],dat1[,6],paired = TRUE)$p.value
    sig<-ifelse(pval<0.05|w.pval<0.05,"*","")
    
    res1<-c(levels(dat$assay)[i],media,colMeans(dat1[,3:6]),pval,w.pval,sig)
    res<-rbind(res,res1)
    
  }
  colnames(res)[7]<-"t.test"
  colnames(res)[8]<-"wil.test"
  
  colnames(res)[1:2]<-c("assay","media")
  row.names(res)[1:6]<-1:6
  
  res<-as.data.frame(res)
  
  res1<-apply(res[,3:8],2,function(x){
    signif(as.numeric(as.character(x)),digits=3)
  })
  
  res2<-cbind.data.frame(res[,1:2],res1,res[,9])
  colnames(res2)[9]<-"sig"
  
return(res2)
  }
```


```{r echo=FALSE}


errort.bars3<-function(gcobj,dbl_strain) {
  par(mfrow=c(3,2))
  
  if(dbl_strain=="CM"){strains<-c("AM373","CssS","McpA","CM","CM_est")
  }else if(dbl_strain=="CN"){strains<-c("AM373","CssS","NhaK","CN","CN_est")
  }else if(dbl_strain=="MN"){strains<-c("AM373","McpA","NhaK","MN","MN_est")}
  
  gcobj2<-gcobj[which(gcobj$strain %in% strains),]
  
  gcobj2$strain<-droplevels(gcobj2$strain)
  gcobj2$strain<-factor(gcobj2$strain,levels=strains,ordered=TRUE)
  gcobj2<-gcobj2[order(gcobj2$strain,gcobj2$plate),]
  
  source('~/Bacillus_subtilis/Growth Rates/growth-curves-ryan/tplot.R')
  for (i in 3:ncol(gcobj2)) {
    
tplot(as.numeric(gcobj2[,i]) ~ gcobj2$strain, data=gcobj2, las=1, cex=1, cex.axis=1, bty='L', show.n=F, dist=.25, jit=.05,
      type=rep('db',times=5),group.pch=T, pch= rep(6,times=length(levels(gcobj2$strain))),
      col=c(1:length(levels(gcobj2$plate)))[gcobj2$plate], boxcol=c(rep('lightsteelblue1',times=3),'blue','coral'),
      boxborder=grey(.8),boxplot.pars=list(notch=F, boxwex=.5),main=colnames(gcobj2)[i])
     abline(h=0,col="red")}
}

```

  

```{r echo=FALSE,warning=FALSE}

#linear modeling
#create version of nrm_a without dbl_est values
nrm3<-nrm_a[-which(grepl("_est",nrm_a$strain)),]

#add strain factor columns - create nrm4
cssS<-relevel(as.factor(ifelse(nrm3$strain %in% c("CssS","CM","CN"),"C","a")),ref="a")
mcpA<-relevel(as.factor(ifelse(nrm3$strain %in% c("McpA","CM","MN"),"M","a")),ref="a")
nhaK<-relevel(as.factor(ifelse(nrm3$strain %in% c("NhaK","CN","MN"),"N","a")),ref="a")

nrm4<-cbind(nrm3[,1:2],cssS,mcpA,nhaK,nrm3[3:ncol(nrm3)])

```


##CM:
```{r warning=FALSE,echo=FALSE,fig.height=8,fig.width=11,width=120}

#select only rows specific to CM-contrast:
cm_data<-nrm4[which(nrm4$strain %in% c("AM373","CssS","McpA","CM")),]

#Run lm on all columns, select those with evidence of epistasis
datl<-NULL
epifnd<-NULL
for(i in 6:ncol(cm_data)){
  datl[[i]]<-summary(lm(cm_data[,i]~cm_data$cssS*cm_data$mcpA))
  if(datl[[i]]$coefficients[4,4]<0.05)
  {epifnd<-rbind(epifnd,c("col"=i, "meas"= colnames(cm_data)[i],                                          "AM373"=signif(datl[[i]]$coefficients[1,1],digits=2), 
        "CssS"=signif(datl[[i]]$coefficients[2,1],digits=2), 
        "McpA"=signif(datl[[i]]$coefficients[3,1],digits=2),
        "CM"= signif(datl[[i]]$coefficients[4,1],digits=2),
        "p-CMinter"=signif(datl[[i]]$coefficients[4,4],digits=2)))}
  }

errort.bars3(nrm_a,"CM")


print("Linear Modeling Results:")
epifnd<-as.data.frame(epifnd)
if(nrow(epifnd)==0){print("No epistatic interactions found by lin mod for CM")}else{epifnd}

print("Estimated Phenotype Calculation Results:")
epi_func(nrm_s,"CssS","McpA","CM","CM_est",media)


```



##CN:
```{r warning=FALSE,echo=FALSE,fig.height=8,fig.width=11,width=120}

#select only rows specific to CM-contrast:
cn_data<-nrm4[which(nrm4$strain %in% c("AM373","CssS","NhaK","CN")),]

#Run lm on all columns, select those with evidence of epistasis
datl<-NULL
epifnd<-NULL
for(i in 6:ncol(cn_data)){
  datl[[i]]<-summary(lm(cn_data[,i]~cn_data$cssS*cn_data$nhaK))
  if(datl[[i]]$coefficients[4,4]<0.05)
  {epifnd<-rbind(epifnd,c("col"=i, "meas"= colnames(cn_data)[i],                                          "AM373"=signif(datl[[i]]$coefficients[1,1],digits=2), 
        "CssS"=signif(datl[[i]]$coefficients[2,1],digits=2), 
        "NhaK"=signif(datl[[i]]$coefficients[3,1],digits=2),
        "CN"= signif(datl[[i]]$coefficients[4,1],digits=2),
        "p-CNinter"=signif(datl[[i]]$coefficients[4,4],digits=2)))}
  }

errort.bars3(nrm_a,"CN")

epifnd<-as.data.frame(epifnd)
if(nrow(epifnd)==0){print("No epistatic interactions found by lin mod for CN")}else{epifnd}

epi_func(nrm_s,"CssS","NhaK","CN","CN_est",media)

```


##MN:
```{r warning=FALSE,echo=FALSE,fig.height=8,fig.width=11,width=120}

#select only rows specific to CM-contrast:
mn_data<-nrm4[which(nrm4$strain %in% c("AM373","McpA","NhaK","MN")),]

#Run lm on all columns, select those with evidence of epistasis
datl<-NULL
epifnd<-NULL
for(i in 6:ncol(mn_data)){
  datl[[i]]<-summary(lm(mn_data[,i]~mn_data$mcpA*mn_data$nhaK))
  if(datl[[i]]$coefficients[4,4]<0.05)
  {epifnd<-rbind(epifnd,c("col"=i, "meas"= colnames(mn_data)[i],                                          "AM373"=signif(datl[[i]]$coefficients[1,1],digits=2), 
        "McpA"=signif(datl[[i]]$coefficients[2,1],digits=2), 
        "NhaK"=signif(datl[[i]]$coefficients[3,1],digits=2),
        "MN"= signif(datl[[i]]$coefficients[4,1],digits=2),
        "p-MNinter"=signif(datl[[i]]$coefficients[4,4],digits=2)))}
  }


  
errort.bars3(nrm_a,"MN")

epifnd<-as.data.frame(epifnd)
if(nrow(epifnd)==0){print("No epistatic interactions found by lin mod for MN")}else{epifnd}


epi_func(nrm_s,"McpA","NhaK","MN","MN_est",media)

```


###Summary:

Significant evidence of epistasis found in all three double-mutants, across multiple growth curve parameters.  Only parameter without apparent epistasis in any strain is ln_mu.spline (max GR in ln-transformed growth curve)