---
title: GC analysis - CssS set on acidic (pH5) LB
author: "Kenyon"
date: "December 22, 2015"
output: html_document
---


#### CssS-7 strain set grown on LB adjusted to pH5 for 20 hours;  7 replicates grown

* Import files

Printout of all .csv files containing the relevant growth curve data:

```{r echo=FALSE}
source('~/Bacillus_subtilis/Growth Rates/growth-curves-ryan/GrowthRateScripts_Dec2015/1_import_csvs.R')

folder_name<-"C:/Users/Kenyon/Documents/Bacillus_subtilis/Growth Rates/Eon data/LB+HCL-PH5/" 
file_names<-list.files(folder_name,pattern="*csv") #Fill me in
file_names<-file_names[grep("20hrs",file_names)]
media<-"LB_pH5"  #fill me in
today<-"2016-06-14"

obj_name<-"csvob"

file_names

```

* Merge individual '.csv' files into one:

```{r echo=FALSE}
for (i in 1:length(file_names)){
  x<-read.csv(paste0(folder_name,file_names[i]),header = FALSE)
  assign(paste0(obj_name,i),x)
}

#Check to make sure length of this is correct:
csvob<-rbind(csvob1[,1:125],csvob2[-1,1:125],csvob3[-1,1:125],csvob4[-1,1:125],csvob5[-1,1:125],csvob6[-1,1:125],csvob7[-1,1:125])

#csvob[,2]<-c("media",rep(media,times=nrow(csvob)-1))

```

* Correct for plate reader OD saturation:

```{r echo=FALSE}

nums<-csvob[-1,-1:-4]

csvob2<-(0.622*nums*nums)+(0.77*nums)

csvob3<-rbind(csvob[1,-1:-4],csvob2)
csvob<-cbind(csvob[,1:4],csvob3)

write.csv(csvob,paste0(folder_name,media,"_rawdatacollected_ODcorrected_",today,".csv"))

```

* Preliminary run of data to check for correct feature identification

```{r eval=FALSE,echo=FALSE,width=120}
#open grofit function, no overlay of plots so bad plots can be identified

source('~/Bacillus_subtilis/Growth Rates/growth-curves-ryan/no_overlay-gc_combined/grofit-combined-no_overlay.R')
par(mfrow=c(1,1))
gcdat<-gcfall_no(csvob)   #already run, do not need to do again

```


####observations:
All plots appear fitted correctly


* apply version of 'gcfall' that merges ln=TRUE and ln=FALSE datasets

###Growth Curve Plots:  

(dashed line is ploted on each curve to highlight the position and slope of the identified max GR)  

```{r echo=FALSE,results='hide',message=FALSE,width=120,warning=FALSE}
source('~/Bacillus_subtilis/Growth Rates/growth-curves-ryan/GrowthRateScripts_Dec2015/3_gcfall-media-lnmerge_lnlag.R')

gcdat2<-gcfall_mergeln(csvob)
```

Plots of each mutant strains (color) relative to AM373 (grey):
    - there is one plot per well grown in these images, so 7 experiments x #technical replicates/experiments 

```{r echo=FALSE,fig.height=12}

par(mfrow=c(3,2))

st.levs<-c("AM373","CssS","McpA","NhaK","CM","CN","MN")
col.sel<-c("black","blue","red","orange","darkorchid","green4","deeppink","lawngreen","hotpink","darkslategrey","deeppink4")

ymax<-round(max(csvob[-1,-c(1:5)])*1.1,digits=1)

for(j in 2:7){
  
c1<-csvob[c(1,which(csvob[,3] %in% c("AM373",st.levs[j]))),]

plot(as.numeric(c1[2,5:ncol(c1)])~as.numeric(c1[1,5:ncol(c1)]),type="l",
     xlab="hours",ylab="OD(600nm)-well",col=ifelse(c1[2,3]=="AM373","gray44",
     col.sel[j]),main=st.levs[j],ylim=c(0,ymax))

for(i in 3:nrow(c1)){
  lines(as.numeric(c1[1,5:ncol(c1)]),as.numeric(c1[i,5:ncol(c1)]),col=ifelse(c1[i,3]=="AM373","gray44",col.sel[j]))
  
}
}
```


#### save data and combine technical replicates:

Each experiment had 2-4 technical replicates in which the same strain was grown in multiple wells; this function calculates the average value of each feature found among the technical replicates.
```{r width=120}
write.csv(gcdat2,paste0(folder_name,media,"-gcfit-",today,".csv"))

# Combine technical replicates
source('~/Bacillus_subtilis/Growth Rates/growth-curves-ryan/LM_controls-Sept2015/average_triples_ingcobj.R')

gcavg<-avg_triples(gcdat2,"DEF")

write.csv(gcavg,paste0(folder_name,media,"-gcfit-repsavgd",today,".csv"))

```


### plot of GR paramaters  

* ln_mu.spline :  max growth rate of ln-transformed data (doublings/hr)  
* mu.spline: max growth rate, raw data (delta-OD/hr)  
* lambda.spline:  lag time (hrs)  
* A.spline: MaxOD of growth curve  (OD)  
* integral.spline: integrated area under each growth curve  


```{r warning=FALSE,fig.width=11,fig.height=9}
#source('~/Bacillus_subtilis/Growth Rates/growth-curves-ryan/LM_controls-Sept2015/gcf-tplot-by-strain.R')

gcavg$strain<-factor(gcavg$strain,levels=c("AM373","CssS","McpA","NhaK","CM","CN","MN"))

source('~/Bacillus_subtilis/Growth Rates/growth-curves-ryan/LM_controls-Sept2015/gcf-tplot-by-strain2.R')
errort.bars(gcavg)

```

##Table of GR parameters

avg: average of all replicates
sd:  sd across replicates
pval: results of paired t-test against AM373 reference strain values (no Mult Comp correction)
sig: if <0.05, marked with "*"
%AM373: percent differenc from AM373


```{r echo=FALSE,results='asis'}

library(knitr)
for(i in 11:14){
tab1<-tapply(gcavg[,i],gcavg$strain,mean)
tab2<-tapply(gcavg[,i],gcavg$strain,sd)
tab3<-tapply(gcavg[,i],gcavg$strain,function(x){
  
  t.test(x,gcavg[which(gcavg$strain=="AM373"),i],paired=TRUE)$p.val
  
})
tab4<-ifelse(tab3<0.05,"*"," ")
tab5<-(tab1-tab1[1])/tab1[1]
tab<-rbind.data.frame(signif(tab1,digits=3),signif(tab2,digits=3),signif(tab3,digits=3),tab4,signif(tab5,digits=3))

lab<-c("avg","sd","pval","sig","%AM373")
lab2<-paste(unlist(strsplit(colnames(gcavg)[i],"[.]"))[1],lab)
rownames(tab)<-lab2
colnames(tab)<-names(tab1)

print(kable(tab,format="markdown",digits=3))
  }

```


**lambda.spline:**

Only MN does not have significantly lower lag than AM373

**ln_mu.spline:**

Looks interesting, but unfortunately, none of the variation is stat sig


**A.spline:**

All 3 double-mutants have sig lower maxOD than AM373 (and McpA, NhaK, and CssS)
MN is sig lower than all strains
    
### Tukey HSD analysis

Identify pairs of strains that are significantly different from each other in each growth feature  

Only pairs that are significantly different (p<0.05): 

```{r width=120}
source('~/Bacillus_subtilis/Growth Rates/growth-curves-ryan/LM_controls-Sept2015/gcftukeyrun_rM.R')

tuk_dat<-tukeyrun(gcavg,c(11,13:14))

tuk_dat[,3:5]<-signif(tuk_dat[,3:5],digits = 3)

print(tuk_dat[which(tuk_dat$pvals<0.05),])

```

All pairs:

```{r width=120}
print(tuk_dat)
```