---
title: GC analysis - template
author: "Kenyon"
date: "December 22, 2015"
output: html_document
---


#### CssS-7 strain set grown on LB adjusted to pH5 for 20 hours;  7 replicates grown

* Import files

```{r echo=FALSE}
source('~/Bacillus_subtilis/Growth Rates/growth-curves-ryan/GrowthRateScripts_Dec2015/1_import_csvs.R')

folder_name<-"C:/Users/Kenyon/Documents/Bacillus_subtilis/Growth Rates/Eon data/LB+HCL-PH5/" 
file_names<-list.files(folder_name,pattern="*csv") #Fill me in
file_names<-file_names[grep("20hrs",file_names)]
media<-"LB_pH5"  #fill me in
today<-"2016-02-10"

obj_name<-"csvob"

file_names

```

* Merge individual '.csv' files into one:

```{r echo=FALSE}
for (i in 1:length(file_names)){
  x<-read.csv(paste0(folder_name,file_names[i]),header = FALSE)
  assign(paste0(obj_name,i),x)
}

#Check to make sure length of this is correct:
csvob<-rbind(csvob1,csvob2[-1,],csvob3[-1,],csvob4[-1,],csvob5[-1,],csvob6[-1,],csvob7[-1,])

#csvob[,2]<-c("media",rep(media,times=nrow(csvob)-1))

```

*account for OD saturation:

```{r echo=FALSE}

nums<-csvob[-1,-1:-4]

csvob2<-(0.622*nums*nums)+(0.77*nums)

csvob3<-rbind(csvob[1,-1:-4],csvob2)
csvob<-cbind(csvob[,1:4],csvob3)

write.csv(csvob,paste0(folder_name,media,"_rawdatacollected_ODcorrected_",today,".csv"))

```

* Preliminary run of data to check for correct feature identification

```{r eval=FALSE,echo=FALSE,width=120}
#open grofit function, no overlay of plots so bad plots can be identified

source('~/Bacillus_subtilis/Growth Rates/growth-curves-ryan/no_overlay-gc_combined/grofit-combined-no_overlay.R')
par(mfrow=c(1,1))
gcdat<-gcfall_no(csvob)   #already run, do not need to do again

```


####observations:
All plots look fine


* apply version of 'gcfall' that merges ln=TRUE and ln=FALSE datasets

###Growth Curve Plots:

```{r echo=FALSE,results='hide',message=FALSE,width=120,warning=FALSE}
source('~/Bacillus_subtilis/Growth Rates/growth-curves-ryan/GrowthRateScripts_Dec2015/3_gcfall-media-lnmerge.R')

gcdat2<-gcfall_mergeln(csvob)
```

**observations:**


#### save data and combine technical replicates:
```{r width=120}
write.csv(gcdat2,paste0(folder_name,media,"-gcfit-",today,".csv"))

# Combine technical replicates
source('~/Bacillus_subtilis/Growth Rates/growth-curves-ryan/LM_controls-Sept2015/average_triples_ingcobj.R')

gcavg<-avg_triples(gcdat2,"DEF")

write.csv(gcavg,paste0(folder_name,media,"-gcfit-repsavgd",today,".csv"))

```


### plot of GR paramaters

```{r warning=FALSE,fig.width=11,fig.height=9}
source('~/Bacillus_subtilis/Growth Rates/growth-curves-ryan/LM_controls-Sept2015/gcf-tplot-by-strain.R')
errort.bars(gcavg)

```


**lambda.spline:**

Only MN does not have significantly lower lag than AM373

**ln_mu.spline:**

Looks interesting, but unfortunately, none of the variation is stat sig


**A.spline:**

All 3 double-mutants have sig lower maxOD than AM373 (and McpA, NhaK, and CssS)
MN is sig lower than all strains
    
```{r width=120}
source('~/Bacillus_subtilis/Growth Rates/growth-curves-ryan/LM_controls-Sept2015/gcftukeyrun_rM.R')

tuk_dat<-tukeyrun(gcavg,c(11,13:14))

tuk_dat[,3:5]<-signif(tuk_dat[,3:5],digits = 3)

print(tuk_dat[which(tuk_dat$pvals<0.05),])
print(tuk_dat)
```



